### 2019-08-21 tsibble.Rmd
done with conclusion revision

---
title: "A new tidy data structure to support exploration and modeling of temporal data"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Dianne
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
abstract: 'Mining temporal data for information is often inhibited by a multitude of formats: irregular or multiple time intervals, point events that need aggregating, multiple observational units or repeated measurements on multiple individuals, and heterogeneous data types. On the other hand, the software supporting time series modeling and forecasting, makes strict assumptions on the data to be provided, typically requiring a matrix of numerical data with implicit time indexes. Going from raw data to model-ready data is painful. This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization, modeling and forecasting routines. Tidy data principles are extended to temporal data by: (1) mapping the semantics of a dataset into its physical layout; (2) including an explicitly declared index variable representing time; (3) incorporating a "key" comprising single or multiple variables to uniquely identify units over time. This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based contexts. A sound data pipeline facilitates a fluent workflow for analyzing temporal data. The infrastructure of tidy temporal data has been implemented in the R package **tsibble**.'
keywords: "time series, data wrangling, tidy data, R, forecasting, data science, exploratory data analysis, data pipelines"
wpnumber: 12/19
jelcodes: C88, C81, C82, C22, C32
blind: false
cover: false
toc: false
bibliography: [references.bib]
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    dev: "pdf"
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
  cache = TRUE, cache.path = 'cache/', 
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

```{r external, include = FALSE, cache = FALSE}
read_chunk('scripts/demo.R')
read_chunk('scripts/flights.R')
read_chunk('scripts/smart-meter.R')
```

```{r load-pkg}
```

# Introduction {#sec:intro}

Temporal data arrives in many possible formats, with many different time contexts. For example, time can have various resolutions (hours, minutes, and seconds), and can be associated with different time zones with possible adjustments such as daylight saving time. Time can be regular (such as quarterly economic data or daily weather data), or irregular (such as patient visits to a doctor's office). Temporal data also often contains rich information: multiple observational units of different time lengths, multiple and heterogeneous measured variables, and multiple grouping factors. Temporal data may comprise the occurrence of time-stamped events, such as flight departures.

Despite this variety and heterogeneity, little organization or conceptual oversight on how one should get the wild data into a tamed state is provided for temporal data. Analysts are expected to do their own data preprocessing and take care of anything else needed to allow further analysis, which leads to a myriad of ad hoc solutions and duplicated efforts.

@r4ds proposed the tidy data workflow, which provides a conceptual framework for exploring data (as described in Figure \@ref(fig:tsibble-fit)). In temporal domain, data with time information arrives at the "import" stage. A new abstraction, *tsibble*, is the gatekeeper at the "tidy" stage, to verify if the raw temporal data is appropriate for downstream analytics. The exploration loop will be aided with declarative grammars, yielding more robust and accurate analyses.

```{r tsibble-fit, fig.cap = "(ref:tsibble-fit-cap)"}
include_graphics("img/tsibble-fit.png")
```

(ref:tsibble-fit-cap) Illustration of the data science workflow regarding temporal data, drawn from @r4ds. The new data structure, tsibble, makes the connection between temporal data input, and downstream analytics. It provides elements at the "tidy" step, which produce tidy temporal data for temporal visualization and modeling.

The paper is structured as follows. Section \@ref(sec:str) reviews temporal data structures corresponding to time series and longitudinal analysis, and discusses "tidy data" and the grammar of data manipulation. Section \@ref(sec:semantics) proposes contextual semantics for temporal data, built on top of tidy data. The concept of data pipelines with respect to the time domain will be discussed in depth in Section \@ref(sec:pipeline), followed by a discussion of the design choices made in the software structure in Section \@ref(sec:software). Two case studies are presented in Section \@ref(sec:cases) illustrating temporal data exploration using the newly implemented infrastructure. Section \@ref(sec:conclusion) discusses future work.

# Data structures {#sec:str}

## Time series and longitudinal data {#sec:time-series}

Temporal data problems are typically grouped into two types of analysis, time series and longitudinal. Despite being exactly the same data input, the representation of time series and longitudinal data diverges due to different modeling focuses.

Time series can be univariate or multivariate, and for modeling require relatively long lengths (i.e., large $T$). Time series researchers and analysts who are concerned with this large $T$ property, are mostly concerned with stochastic processes, for the primary purpose of forecasting, and characterizing temporal dynamics. Most statistical software represent such time series as vectors or matrices. Multivariate time series are typically assumed to be in the format where each row is assumed to hold observations at a time point and each column to contain a single time series. (The tidy data name for this would be **wider format**.) This implies that data are columns of homogeneous types: numeric or non-numeric, but there are limited supporting methods for non-numeric vectors. Time indices are implicitly inferred as attributes or meta-information. The number of observations must be the same across all the series. Data wrangling, from the form that data arrives in to this specialist format, can be mysterious and frustrating. This kind of format is model-centric, and the diagram in the style of Figure \@ref(fig:tsibble-fit) would place the "model" at the center of the analytical universe, and all the transformations and visualizations would hinge on that format. This is contrary to the **tidyverse** conceptualization, which holistically captures the full data workflow.

For longitudinal analysis, researchers and analysts are primarily interested in explaining trends across and variations among individuals, and making inference about a broader population. Longitudinal data or panel data typically assumes fewer measurements (small $T$) over a large number of individuals (large $N$). It often occurs that measurements for individuals are taken at different time points. Thus, the primary format required for modeling such data is stacked data, blocks of measurements for each individual, with columns indicating panels, times of measurement and the measurements themselves. (The tidy data name for this would be **longer format**.) Evidently, this data organization saves storage space for many sparse cells, compared to structuring it into the wider format which would have missing values in many cells. An appealing feature is that data is structured in a semantic manner with reference to observations and variables, with panel and time stated explicitly. It is conducive to examining the data in many different ways and leading to more comprehensive exploration.

## Existing time series and longitudinal data standards

In R [@R-base], time series and longitudinal data are of different representations. The native `ts` object and the enhancement of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wider matrices with implicit time indices. A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent time series in heterogeneous longer format. It only requires an index variable to be declared. However, as proposed in Section \@ref(sec:semantics) a more rigid data structure is required for temporal analytics and models. The **plm** [@plm2008] and **panelr** [@R-panelr] packages both manage longitudinal data in terms of longer format.

Stata [@Stata] provides two commands, `tsset` and `xtset`, to declare time series and panel data respectively, both of which need specifying panel and time variables. Note they are in fact the same command, according to the manual. The underlying data arrangement is the unifying longer form. Either time series or panel commands are sufficient to apply to both types. The SAS software [@SAS-STAT] handles them in a similar way as Stata.

## Tidy data and the grammar of data manipulation

@wickham2014tidy coined the term "tidy data", which is a rephrasing of the second and third normal forms in relational databases but in a way that makes more sense to data scientists by referring rows to observations and columns to variables. The principles of "tidy data" attempt to standardize the mapping of the semantics of a dataset to its physical representation. This data structure is the fundamental unit of the **tidyverse**, which is a collection of R packages designed for data science. The ubiquitous use of the **tidyverse** is testament to the simplicity, practicality and general applicability of the tools. The **tidyverse** provides abstract yet functional grammars to manipulate and visualize data in easier-to-comprehend form. One of the **tidyverse** packages, **dplyr** [@R-dplyr], showcases the value of a grammar as a principled vehicle to transform data for a wide range of data challenges, providing a consistent set of verbs: `mutate()`, `select()`, `filter()`, `summarize()`, and `arrange()`. Each verb focuses on a singular task. Most common data tasks can be rephrased and tackled with these five key verbs, in conjunction with `group_by()` to perform grouped operations.

The **tidyverse** largely formalizes exploratory data analysis. Many in the R community have adopted the **tidyverse** way of thinking and extended it to broader domains, such as simple features for spatial data in the **sf** package [@RJ-2018-009] and missing value handling in the **naniar** package [@tierney-naniar-2018]. This paper with the associated **tsibble** R package [@R-tsibble] extends the tidy way of thinking to temporal data.

This paper attempts to provide a definition of tidy temporal data in relation to "tidy data", which encompasses both time series and longitudinal data. It aims at a unified framework to streamline the workflow from data preprocessing to visualization and modeling, as an integral part of a tidy data analysis.

# Contextual semantics {#sec:semantics}

The choice of tidy representation of temporal data arises from a data- and model-oriented perspective, which can accommodate all of the operations that are to be performed on the data in time-based contexts. Figure \@ref(fig:tsibble-fit) marks where this new abstraction is placed in the "tidy" model, which is referred to as a "tsibble". Adhering to the "tidy data" principles, in tsibble:

1. Index is a variable with inherent ordering from past to present.
2. Key is a set of variables that define observational units over time.
3. Each observation should be uniquely identified by index and key.
4. Each observational unit should be measured at a common interval, if regularly spaced.

Figure \@ref(fig:tsibble-semantics) sketches out the data form required for a tsibble, an extension of the "tidy" format to the time domain. Beyond the layout, tsibble gives the contextual meaning to variables in order to construct the temporal data object, as newly introduced "index" and "key" semantics stated in the definitions 1 and 2 above. Variables other than "index" and "key" are considered as measurements. The definitions of 3 and 4 imply that tsibble is *tidier* than "tidy data", positioning itself as a model input that gives rise to more robust and reliable downstream analytics.

```{r tsibble-semantics, fig.cap = "(ref:tsibble-semantics-cap)"}
include_graphics("img/tsibble-semantics.png")
```

(ref:tsibble-semantics-cap) The architecture of the tsibble structure is built on top of the "tidy data" principles, with temporal semantics: index and key.

To materialize the abstraction of the tsibble, a subset of tuberculosis cases estimated by @tb-data presented in Table \@ref(tab:tb-sub) is used as an example. It contains 12 observations and 5 variables landing in a tidy data form. Each observation comprises the number of people who are diagnosed with tuberculosis for each gender at three selected countries in the years of 2011 and 2012. From tidy data to tsibble data, "index" and "key" should be declared: column `year` as the "index" variable, and column `country` together with `gender` as the "key" variable forming the observational units. Column `count` is the only measured variable in this data, but the data structure is sufficiently flexible to hold more measurements; for example, slotting the corresponding population size (if known) into the data column for normalizing the count later. Note this data further satisfies the must-have for the distinct rows determined by index and key, and regularly spans over one-year interval.

```{r tb-sub}
```

(ref:tb-sub) A small subset of estimates of tuberculosis burden collected by World Health Organization in 2011 and 2012, with 12 observations and 5 variables. The index refers to column `year`, the key to multiple columns: `country` and `gender`, and the measured variable to column `count`.

The new tsibble structure bridges the gap between raw data and the rigorous state of temporal data analysis. The proposed contextual semantics is the new add-on to tidy data in order to support more intuitive time-related manipulations and enlighten new perspectives for time series and panel model inputs. Index, key and time interval build the three stone pillars to this new semantically structured temporal data. Each is now described in more detail.

## Index {#sec:index}

> *Index is a variable with inherent ordering from past to present.*

Time provides the contextual basis for temporal data. Time can be seen in numerous representations, from sequential numerics to the most commonly accepted date-times. Regardless of this diversity, time should be inherently ordered from past to present, so should the index variable to a tsibble.

Index is an explicit data variable rather than a masked attribute (such as in the `ts` and `zoo` classes), exposing a need for more accessible and transparent time operations. It is often to visualize and model seasonal effects of measurements of interest, meaning that time components such as time of day and day of week from the index should be easily extracted upon request. But when the index is available only as meta information, an obstacle is inevitably created for analysts to write such simple queries in an ad hoc and programmatic manner. From an analytical point of view it should be discouraged.

## Key {#sec:key}

> *Key is a set of variables that define observational units over time.*

What subjects/entities to be observed over time leads to the second component of a tsibble--key. Key can consist of empty, single, or multiple variables identifying units measured along the way. When only a single observational unit is present in the table, key can be left with no variables to be specified. However, when multiple units exist in the data, key should be supplied by identifying variables to sufficiently define the units. In longitudinal data, key can be thought of as "panel" (such as in the Stata) but constrained to a single variable in existing data structures. In tsibble, key allows for multiple variables of nesting, crossing, or union relations [@wilkinson2006grammar], that can be useful for forecasting reconciliation [@fpp; @R-hts] and richer visualization. For example, Table \@ref(tab:tb-sub) describes the number of tuberculosis cases for each gender across the countries every year. This suggests that the key comprises at least columns `gender` and `country`. Since `country` is nested within `continent`, it is a free variable to be included in the key specification.

> *Each observation should be uniquely identified by index and key.*

Inspired by "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, tsibble key also uniquely identifies each observational unit over time. When constructing a tsibble, any duplicates of key-index (i.e. key-value) pairs will fail, because duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making. For example, either `gender` or `country` alone is not enough to be the key for the tuberculosis data. Analysts are encouraged to better understand the data, or reason about the process of data cleaning when handling duplicates. Figure \@ref(fig:tsibble-tidy) peels the "tidy" module with clear routes required for a tsibble. The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage.

```{r tsibble-tidy, fig.cap = "(ref:tsibble-tidy-cap)"}
include_graphics("img/tsibble-tidy.png")
```

(ref:tsibble-tidy-cap) Details about the "tidy" stage for a tsibble. Built on top of "tidy data", each observation should be uniquely identified by index and key, thereby no duplicated key-index pairs.

Since observational units are embedded, modelling and forecasting for each unit in a tsibble will be accelerated. The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts. This neatly decouples the expensive data copy from downstream summaries, significantly reducing the storage space.

## Interval {#sec:interval}

> *Each observational unit should be measured at a common interval, if regularly spaced.*

The principal divide of temporal data is regularly versus irregularly spaced data. Event data typically involves irregular time intervals, such as flight schedules or customer transactions. This type of data can flow into event-based data modeling, but would need to be processed or regularized to fit models that expect data with a fixed-time interval. When data indexed in regular time space, tsibble computes the greatest common divisor as the time interval from the index, or an unknown interval if one observation is available for each unit.

By extending one of the "tidy data" rules, "Each type of observational units should form a table", each observational unit shares a common interval in a tsibble. Although tsibble reports one single interval despite data of fixed or mixed intervals, data of mixed intervals should be organized in separate tsibbles instead for well-tailored analysis and models.

This tiny piece of information will be carried over for tsibble-centric functions, for example to make implicit missing time handling handy and be in harmony with statistics and models that use seasonal periods.

# Temporal data pipelines {#sec:pipeline}

A data pipeline describes the flow of data through an analysis, and can generally assist in conceptualizing the process for a stream of problems. @unix coined the term "pipelines" in software development while developing Unix at Bell Labs. In Unix-based computer operating systems, a pipeline chains together a series of operations based on their standard streams, so that the output of each program becomes the input to another. The Extract, Transform, and Load (ETL) process from recent data warehousing literature dating back to @kimball2011data outlines the workflow to prepare data for analysis, and can also be considered a data pipeline. @viewing-pipeline describes a viewing pipeline for interactive statistical graphics, that takes control of the transformation from data to plot. @xgobi, @ggobi, @orca, @plumbing and @xie2014reactive implemented data pipelines for the interactive statistical software **XGobi**, **GGobi**, **Orca**, **plumbr** and **cranvas**, respectively.

<!--
The pipeline is typically described with a one way flow, from data to plot. For interactive graphics, where all plots need to be updated when a user interacts with one plot, the events typically trigger the data pipeline to be run. @xie2014reactive uses a reactive programming framework, to implement the pipeline, in which user's interactions trigger a sequence of modules to update their views, that is, practically the same as running the data pipeline producing each plot.
-->

A fluent data pipeline anticipates a standard data structure. The tsibble data abstraction lays the plumbing for data analysis modules of transformation, visualization and modeling in temporal contexts. It provides a data infrastructure to a new ecosystem, **tidyverts**, a play on tidyverse that acknowledges the time series analysis purpose, in the spirit of "tidy data" to the **tidyverse**.

## Transformation {#sec:timepipeline}

The **tsibble** package not only provides a "tsibble" data object but also a domain specific language in R for transforming temporal data. It takes advantage of the wrangling verbs implemented in the **dplyr** package, and develops a suite of new tools for facilitating temporal manipulation with ease primarily in two aspects: implicit missingness handlers and time-aware aggregations.

Implicit missingness are values that should be present but are absent. Regarding regularly-spaced temporal data, data entries that should be available at certain time points but are missing, leaving gaps in time. It will be a problem when temporal models and operations like lag/lead are applied. A family of verbs help to explore implicit missing values and turn them into the state of explicitness to eliminate potential concerns, listed as follows:

* `has_gaps()` checks the existence of time gaps.
* `scan_gaps()` reveals all implicit missing observations.
* `count_gaps()` summarizes the time ranges that are absent from the data.
* `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.

These verbs are evocative and of simple interface. They by default look into gaps for each individual time period. Switching on the option `.full = TRUE` will fill in the full-length time span, creating fully balanced panels with respect to longitudinal study.

The other key function, an adverb `index_by()`, is the counterpart of `group_by()` in **dplyr**, grouping and partitioning the index only. It is most useful to be used in conjunction with `summarize()`, and thus aggregations to upper-level time resolutions will take place. This combination automatically produces new index and interval, and helps to regularize data of irregular interval.

Besides new verbs, the **dplyr** vocabulary has been adapted and expanded to facilitate temporal transformations. The **dplyr** suite showcases the general-purpose verbs for effectively manipulating tabular data. But these verbs need handling with care due to the context switch. A perceivable difference is summarizing variables between normal data and tsibble using `summarize()`. The former will reduce to a single summary, whereas the latter will obtain the index and their corresponding summaries.

Attention has been paid to warning and error handling. The principle that underpins all the verbs is a tsibble in and a tsibble out, thereby striving to maintain a valid tsibble over the course of transformation. If the desired temporal ordering is changed by row-wise verbs (such as `arrange()` and `slice()`), a warning is broadcast. If a tsibble cannot be maintained in the output of a pipeline module (likely occurring to column-wise verbs), for example the index is dropped by `select()`-ing, an error informs users of the problem and suggests alternatives. This avoids surprising users and reminds them of the time context. In general, users who are already familiar with the **tidyverse**, should have less resistance to learning the new semantics and verbs.

## Visualization {#sec:ts-vis}

The **ggplot2** package (as the implementation of grammar of graphics) builds a powerful graphical system to declaratively visualize data. The data underpinning of **ggplot2** is tidy data, and in turn tsibble integrates well with **ggplot2**. The integration encourages more flexible graphics for exploring temporal structures via index, and individual or group differences via key.

Line charts are universally accepted for ordered data, such as time series plots or spaghetti plots, depending on the fields. But they end up with exactly the same grammar: chronological time mapped to the horizontal axis, and the interested measurement on the vertical axis, for each unit. Many specialist plots centering around time series or longitudinal data, hence can be described and re-created under the umbrella of the grammar and **ggplot2**.

## Model {#sec:ts-models}

Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis start to diverge. Nevertheless, tsibble as a model-oriented object can flow into both types of models, with new semantics (index and key) to be internally utilized.

Most time series models are univariate, such as ARIMA and Exponential Smoothing, modelling temporal dynamics for each series independently. The **fable** package [@R-fable] provides a tidy forecasting framework built on top of tsibble, which is under development, with the goal of promoting transparent and human-centered forecasting practices. With the presence of the key, tsibble holds many series. Since models are fundamentally scalable, the `model()` and `forecast()` generics will take care of fitting and forecasting univariate models to each series across time in a tsibble at once.

Panel data models, however, put emphases on overall, within, and between variations both across individuals and time. Fixed and random effects models can be developed in line with the **fable** design.

## Summary

To sum it up, the tsibble abstraction provides a formal organization of forwarding tidy data to model-oriented temporal data. The supporting operations can be chained for sequencing analysis, articulating a data pipeline. As @eopl stated, "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." A mini snippet below glues transformation and forecasting together, to envision the fluent pipeline.

```{r tsibble-pipeline, eval = FALSE, echo = TRUE}
```

The `pedestrian` dataset [@ped] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package. The pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr] chains the verbs, read as "then". A sequence of functions are composed in a way that can be naturally read from left to right, which improves the code readability.

It coordinates a user's analysis making it cleaner to follow, and permits a wider audience to document the data analysis through code without getting lost in a jungle of computational intricacies. It helps (1) break up a big problem to into manageable blocks, (2) generate human readable analysis workflow, (3) avoid introducing mistakes, at least making it possible to trace them through the pipeline.

# Software structure and design decisions {#sec:software}

The **tsibble** package development follows closely to the **tidyverse** design principles [@tidyverse-principles].

## Data first

The primary force that drives the software's design choices is "data". All functions in the package **tsibble** start with `data` or its variants as the first argument, namely "data first". This lays out a consistent interface and addresses the significance of the data throughout the software.

Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal contexts, particularly useful when handling a large collection of data. The contextual information provided by the `print()` function, shown below from Table \@ref(tab:tb-sub), contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of units. These summaries aid users in understanding their data better.

```{r tb-print}
```

## Functional programming

Rolling window calculations are widely used techniques in time series analysis, and often apply to other applications. These operations are dependent on having an ordering, particularly time ordering for temporal data. Three common types of variations for sliding window operations are:

1. **slide**: sliding window with overlapping observations.
2. **tile**: tiling window without overlapping observations.
3. **stretch**: fixing an initial window and expanding to include more observations.

Figure \@ref(fig:animate) shows the animations of rolling windows for sliding, tiling and stretching, respectively, on annual tuberculosis cases for Australia. A block of consecutive elements with a window size of 5 are initialized and started rolling sequentially till the end of series by computing average counts.

```{r tb-au}
```

```{r animate, fig.show = "animate", fig.height = 7, fig.cap = "(ref:animate-cap)", interval = 0.5, aniopts = c("autoplay", "controls", "loop")}
```

(ref:animate-cap) An illustration of window of size 5 computing rolling averages over annual tuberculosis cases in Australia with respect to sliding, tiling and stretching. (Animation needs to be viewed with Adobe Acrobat Reader.)

Rolling window adapts to functional programming, which the **purrr** package [@R-purrr] sets a good example. These functions accept and return arbitrary inputs and outputs, with arbitrary methods. For example, moving averages anticipate numerics and produce averaged numerics via `mean()`. However, rolling window regression feeds a data frame into a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.

Rolling window not just iterates but rolls over a sequence of elements of a fixed window. A complete and consistent set of tools are available for facilitating window-related operations, a family of `slide()`, `tile()`, `stretch()`, and their variants. `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs. For type stability, the functions always return lists. Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding types, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively. Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel [@R-future; @R-furrr].

## Modularity

Modular programming is adopted in the design of the **tsibble** package. Modularity benefits users by providing small focused and cleaner chunks, and provides developers with simpler maintenance.

All user-facing functions can be roughly organized into three major chunks according to their functionality: vector functions (1d), table verbs (2d), and window family. Each chunk is an independent module, but works interdependently. Vector functions in the package mostly operate on time. The atomic functions (such as `yearmonth()` and `yearquarter()`) can be embedded in the `index_by()` verb to collapse a tsibble to a less granular interval. Since they are not tied to a tsibble, they can be used in a broader range of data applications not constrained to tsibble. On the other hand, the table verbs can incorporate many other vector functions from a third party, like the **lubridate** package.

## Extensibility

As a fundamental infrastructure, extensibility is a design decision that was employed from the start of **tsibble**'s development. Contrary to the "data first" principle for end users, extensibility is developer focused and would be mostly used in dependent packages, which heavily relies on S3 classes and methods in R [@adv-r]. The package can be extended in two major aspects: custom index and new tsibble class.

Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution in **nanotime** [@R-nanotime], and numerics in simulation. Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday. The **tsibble** package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters. These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year. A new index would be immediately recognized upon defining `index_valid()`, as long as it can be ordered from past to future. The interval regarding semesters is further outlined through `interval_pull()`. As a result, all tsibble methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.

The class of tsibble is an underpinning for temporal data, and sub-classing a tsibble will be a demand. A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass. This new object itself is a tsibble. It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, for example prediction distributions to a forecasting tsibble.

## Tidy evaluation

The **tsibble** packages leverages the **tidyverse** grammars and pipelines through tidy evaluation [@tidy-eval] via the **rlang** package [@R-rlang]. In particular, the table verbs extensively use tidy evaluation to evaluate computation in the context of tsibble data and spotlights the "tidy" interface that is compatible with the **tidyverse**. This not only saves a few keystrokes without explicitly repeating references to the data source, but the resulting code is typically cleaner and more expressive, when doing interactive data analysis.

# Case studies {#sec:cases}

## On-time performance for domestic flights in U.S.A

```{r load-flights-pkgs}
```

```{r load-flights, cache.extra = tools::md5sum("data/flights.rds")}
```

The dataset of on-time performance for US domestic flights in 2017 represents event-driven data caught in the wild, sourced from US Bureau of Transportation Statistics [@flights-data]. It contains `r format(NROW(flights), big.mark = ",")` operating flights with many measurements (such as departure delay, arrival delay in minutes, and other performance metrics) and detailed flight information (such as origin, destination, plane number and etc.) in a tabular format. This kind of event describes each flight scheduled for departure at a time point in its local time zone. Every single flight should be uniquely identified by the flight number and its scheduled departure time, from a passenger's point of view. In fact, it fails to pass the tsibble hurdle due to duplicates in the original data. An error is immediately raised when attempting to convert this data into a tsibble, and closer inspection has to be carried out to locate the issue. The **tsibble** package provides tools to easily locate the duplicates in the data with `duplicates()`. The problematic entries are shown below.

```{r find-duplicate}
```

The issue was perhaps introduced when updating or entering the data into a system. The same flight is scheduled at exactly the same time, together with the same performance statistics but different flight details. As flight NK630 is usually scheduled at 17:45 from Chicago to New York (discovered by searching the full database), a decision is made to remove the first row from the duplicated entries before proceeding to the tsibble creation.

```{r tsibble}
```

This dataset is intrinsically heterogeneous, encoded in numbers, strings, and date-times. The tsibble framework, as expected, incorporates this type of data without any loss of data richness and heterogeneity. To declare the flight data as a valid tsibble, column `sched_dep_datetime` is specified as the "index", and column `flight_num` as the "key". As a result of event timing, the data are irregularly spaced, and hence switching to the irregular option is necessary. The software internally validates if the key and index produce distinct rows, and then sorts the key and the index from past to recent. When the tsibble creation is done, the print display is data-oriented and contextually informative, including dimensions, irregular interval (``r tibble::tbl_sum(us_flights)[1]``) and the number of time-based observational units (``r tibble::tbl_sum(us_flights)[2]``).

```{r tsibble-header}
```

Transforming a tsibble for exploratory data analysis with a suite of time-specific and general-purpose manipulation verbs can result in well-constructed pipelines. From the perspective of a passenger, for example, one needs to travel smart by choosing an efficient carrier to fly with and the time of day to avoid congestion. To explore this data, we drill down starting with annual carrier performance and followed by disaggregation to finer time resolutions.

```{r carrier-delayed}
```

```{r carrier-mosaic, fig.cap = "(ref:carrier-mosaic-cap)"}
```

(ref:carrier-mosaic-cap) Mosaic plot showing the association between the size of airline carriers and the delayed proportion of departures in 2017. Southwest Airlines is the largest operator, but does not operate as efficiently as Delta. Hawaiian Airlines, a small operator, outperforms the rest.

```{r flight-pipeline1, fig.cap = "(ref:flight-pipeline1-cap)"}
include_graphics("img/flight-pipeline1")
```

(ref:flight-pipeline1-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:carrier-mosaic).

Figure \@ref(fig:carrier-mosaic) visually presents the end product of aggregating the number of on-time and delayed flights to the year interval by carriers. This pipeline is initialized by defining a new variable if the flight is delayed, and involves summarizing the tallies of on-time and delayed flights for each carrier annually. To prepare the summarized data for a mosaic plot, it is further manipulated by melting new tallies into a single column. The flow chart (Figure \@ref(fig:flight-pipeline1)) demonstrates the operations undertaking in the data pipeline. The input to this pipeline is a tsibble of irregular interval, and the output ends up with a tsibble of unknown interval (as each carrier ends up with only one annual summary). The final data set includes each carrier along with a single year, with the interval undetermined, which in turn feeds into the mosaic display. Note that Southwest Airlines (WN), as the largest carrier, operates less efficiently than Delta (DL).

A closer examination of some big airports across the US will give an indication of how well the busiest airports manage the outflow traffic on a daily basis. A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seattle (SEA) airports is obtained first. The succeeding operations compute delayed percentages every day at each airport, which are framed as grey lines in Figure \@ref(fig:sel-monthly-plot). Winter months tend to fluctuate a lot compared to the summer across all the airports. Superimposed on the plot are two-month moving averages, so the temporal trend is more visible. The number of days for each month is variable. Moving averages for two months call for computing weighted mean. But this can also be accomplished using a pair of commonly used verbs--`nest()` and `unnest()` to handle list-columns, without weight specification. The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year. LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady number delays over time. The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year.

```{r sel-flights}
```

```{r sel-delay}
```

```{r sel-monthly-1}
```

```{r sel-monthly-2}
```

```{r sel-monthly-plot, fig.cap = "(ref:sel-monthly-plot-cap)", fig.height = 7}
```

(ref:sel-monthly-plot-cap) Daily delayed percentages for departure with two-month moving averages overlaid at five international airports. There are least fluctuations and relatively fewer delays observed at KOA airport. The estimates of temporal trend are around 25% across the other four airports, but highlight different time periods of severe delays.

```{r flight-pipeline2, fig.cap = "(ref:flight-pipeline2-cap)"}
include_graphics("img/flight-pipeline2")
```

(ref:flight-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:sel-monthly-plot).

What time of day and day of week should we travel to avoid suffering from horrible delay? Figure \@ref(fig:draw-qtl) plots hourly quantile estimates across day of week in the form of small multiples. The upper-tail delay behaviors are of primary interest, and hence 50%, 80% and 95% quantiles are shown. To reduce the likelihood of suffering a delay, it is recommended to avoid the peak hour around 6pm (18).

```{r quantile}
```

```{r draw-qtl-prep}
```

```{r draw-qtl, fig.cap = "(ref:draw-qtl-cap)"}
```

(ref:draw-qtl-cap) Small multiples of lines about departure delay against time of day, faceting day of week and 50%, 80% and 95% quantiles. A blue horizontal line indicates the 15-minute on-time standard to help grasp the delay severity. Passengers are apt to hold up around 18 during a day, and are recommended to travel early. The variations increase substantially as the upper tails.

```{r flight-pipeline3, fig.cap = "(ref:flight-pipeline3-cap)"}
include_graphics("img/flight-pipeline3")
```

(ref:flight-pipeline3-cap) Flow chart illustrates the pipeline that preprocesses the data for creating Figure \@ref(fig:draw-qtl).

\FloatBarrier

## Smart-grid customer data in Australia

```{r load-elec}
```

Sensors have been installed in households across major cities in Australia to collect data for the smart city project. One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter]. Data from 2013 have been sliced to examine temporal patterns of customer's energy consumption with **tsibble** for this case study. Half-hourly general supply in kwH have been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total. Daily high and low temperatures in Newcastle in 2013 provides explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang]. Two data tables might be joined to explore how local weather can contribute to the variations of daily electricity use when needed.

```{r elec-gaps}
```

During a power outage, electricity usage for some households may become unavailable, thus resulting in implicit missing values in the database. Gaps in time occur to `r scales::percent(sum(gap_df$.gaps) / NROW(gap_df))` of the households in this dataset. It would be interesting to explore these missing patterns as part of a preliminary analysis. Since the smart meters have been installed at different dates for each household, it is reasonable to assume that the records are obtainable for different time lengths for each household. Figure \@ref(fig:count-gaps) shows the gaps for the top 49 households arranged in rows from high to low in tallies. (The remaining households values have been aggregated into a single batch and appear at the top.) Missing values can be seen to occur at any time during the entire span. A small number of customers have undergone energy unavailability in consecutive hours, indicated by a line range in the plot. On the other hand, the majority suffer occasional outages with more frequent occurrence in January.

```{r count-gaps, fig.height = 7, fig.cap = "(ref:count-gaps-cap)"}
```

(ref:count-gaps-cap) Time gap plots for the 49 customers with most implicit missing values, and the remaining customers grouped into the one line at top. Each cross represents an observation missing in time and a line between two dots shows continuous missingness over time. Each row corresponds to one customer. Missing values occur at various times, with more in January and February than other months.

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)). Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees Celsius. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening. This plot illustrates how the tsibble data can easily integrate with other tools and graphics.

```{r calendar-plot, fig.height = 7, fig.cap = "(ref:calendar-plot-cap)"}
```

(ref:calendar-plot-cap) Half-hourly average electricity use across all customers in the region, organized into calendar format, with color indicating hot days. Energy use of hot days tends to be higher, suggesting air conditioner use. Days in the winter months have a double peak suggesting heater use.

```{r elec-pipeline2, fig.cap = "(ref:elec-pipeline2-cap)"}
include_graphics("img/elec-pipeline2")
```

(ref:elec-pipeline2-cap) Flow chart illustrating the pipeline that preprocessed the data for creating Figure \@ref(fig:calendar-plot).

# Conclusion and future work {#sec:conclusion}

A data abstraction, tsibble, is proposed to represent temporal data, allowing the "tidy data" principles to be brought to the time domain. Tidy data begins to take shape in the state of time with new contextual semantics: index and key. The index variable provides direct support to an exhaustive set of ordered objects. The key identifies observational units over time, which can consist of single or multiple variables to be useful in more scenarios. These semantics further determine unique data entries required for a valid tsibble. It shepherds raw temporal data through the "tidy" stage to the next exploration stage for gaining more insights.

The supporting toolkits articulate the temporal data pipeline, with the shared goal of reducing the time between the framings of data questions and the code realization. The rapid iteration for better understanding data is achieved through frictionlessly shifting among transformation, visualization, and modelling, using the standardized tsibble data infrastructure.

A nice feature that the tsibble structure should take into account is to allow user-defined calendars and to respect structurally missing observations. For example, a call center may operate only between 9:00 am and 5:00 pm on week days, and stock trading resumes on Monday straight after Friday. No data available outside trading hours would be labeled as structural missingness, which tsibble currently disregards. Customer calendars can be embedded into the tsibble framework in theory. A few R packages provide functionality to create and manage many specific calendars, such as the **bizdays** package [@R-bizdays] for business days calendars. However, the lack of a generic calendar system delays the implementation. This is left for future work.

# Acknowledgments {-}

The authors would like to thank Mitchell O'Hara-Wild for many discussions on the software development and Davis Vaughan for contributing ideas on rolling window functions. We also thank Stuart Lee for the feedback on this manuscript. We are grateful for anonymous reviewers for helpful feedback that has led to many improvements of the paper. This article was created with knitr [@knitr] and R Markdown [@rmarkdown]. The project's Github repository <https://github.com/earowang/paper-tsibble> houses all materials required to reproduce this article and a history of the changes.

```{r write-bib, eval = FALSE}
write_bib(c("dplyr", "tidyr", "purrr", "nanotime", "tibbletime", "tsibble", "magrittr", "future", "furrr", "bomrang", "rlang", "fable", "panelr", "bizdays"), file = "rpkgs.bib")
```