### 2019-08-30 cover-letter/cover-letter.Rmd
updated cover letter

---
from: Earo Wang
fromqualifications: B.Comm.(Hons)
fromtitle: PhD candidate
fromWWW: earo.me
fromemail: earo.wang@monash.edu
fromphone: +61 4 4919 1031
fromsignature: signature.png
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Sincerely
spacing: 1.4
output: MonashEBSTemplates::letter
---

We have undertaken major revision on our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data," based on reviewers' comments.

We have rewritten Section 3 "Contextual semantics" and Section 4 "Temporal data pipelines". Now Section 3 gives a definition of tsibble in relation to tidy data. Section 4 describes a big picture about each module with the focus of temporal data instead of time series before.

A detailed response letter to reviewers has been provided.

We believe that this manuscript has been significantly improved, that is a good fit for the Journal of Computational and Graphical Statistics.