### 2019-11-10 cover-letter/cover-letter.Rmd
done with minor revisions

---
from: Earo Wang
fromqualifications: B.Comm.(Hons)
fromtitle: PhD candidate
fromWWW: earo.me
fromemail: earo.wang@monash.edu
fromphone: +61 4 4919 1031
fromsignature: signature.png
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Sincerely
spacing: 1.4
output: MonashEBSTemplates::letter
---

We have addressed minor issues that the referees have pointed out on our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data,".

Two zip files are attached with this submission: