### 2019-11-12 cover-letter/cover-letter.Rmd
editted cover letter

---
from: Earo Wang
fromqualifications: B.Comm.(Hons)
fromtitle: PhD candidate
fromWWW: earo.me
fromemail: earo.wang@monash.edu
fromphone: +61 4 4919 1031
fromsignature: signature.png
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Sincerely
spacing: 1.4
output: MonashEBSTemplates::letter
---

We have addressed minor issues that the referees have pointed out on our manuscript titled, "A new tidy data structure to support exploration and modeling of temporal data".

The `source` zip file is attached with this submission, to compile the `.pdf` of the manuscript. This zip file contains:

1. `figure/` Figures generated by R for case studies.
2. `img/` Images generated by Sketch to demonstrate conceptual models.
3. `references.bib` Bibliography items used for BibTeX.
4. `tsibble.tex` LaTeX source file.
5. `tsibble.pdf` A duplicate copy of the non-blinded pdf manuscript file.

The supplemental materials contain the R code used for the case studies, available as the `scripts` zip. We mentioned in the "Acknowledgments" section that all materials contained in the paper can be found at <https://github.com/earowang/paper-tsibble>.